FinalManuscript_DISCS_UG

### 2019-02-14 pcsc-paper/parts/02_01_paillier.tex
Plots for experiment data
Applied most of the reviewer comments
Still missing time benchmarking

\subsubsection{Key Generation}
We first define a function $L\left(x\right)$ as the largest integer $v$ greater than zero such that $x-1 \geq vn$.
We choose two large primes $p$ and $q$, and set $n = pq, \lambda = \mathrm{lcm}\left(p-1,q-1\right)$.
Then we select an integer $g$ such that $\gcd\left(L\left(g^\lambda \bmod n^2\right), n\right) = 1$ and $0\leq g \leq n^2$.
We denote the public key as $(g,n)$ and the private key $(p,q)$.
\subsubsection{Encryption and Decryption}
The encryption function to encrypt a plaintext $m \in \mathbb{Z}_n$ given a public key $(g,n)$ is defined as
\begin{equation*}
	E\left(m\right) = g^m \cdot r^n \mod{n^2},
\end{equation*}
where $r$ is a random non-negative integer less than $n^2$.
The decryption function to decrypt a ciphertext $c \in \mathbb{Z}^*_{n^2}$ given a private key $(p,q)$ is defined as:
\begin{equation*}
	D\left(c\right) = L\left(c^\lambda \bmod n^2\right) \cdot \left(L\left(g^\lambda \bmod n^2\right)\right)^{-1} \mod n.
\end{equation*}
\subsubsection{Homomorphic Properties of the Paillier Cryptosystem}
The Paillier cryptosystem supports additive homomorphism as well as the multiplication of a plaintext scalar to an encrypted message. These homomorphic properties are used to add a plaintext scalar to a ciphertext, add two ciphertexts, and multiply a ciphertext by a plaintext scalar, respectively. They are defined as follows.

@@

We represent a floating-point (FP) number as a pair of two integers $(m,e)$ representing the mantissa and exponent of the FP number with respect to a base $b$. The mantissa $m$ is encrypted, while the exponent $e$ is unencrypted.
Let $a,b,c$ be FP numbers represented by the pairs $(m_a,e_a),(m_b,e_b),(m_c,e_c)$ respectively. Let $\oplus,\otimes$ represent the homomorphic operations which correspond to the addition and multiplication of integers in the Paillier cryptosystem, respectively. We define the corresponding FP number operations as follows:
\begin{description}
  \item[Addition.]
    To compute $E\left(c\right)=E\left(a+b\right)$ we compute
    \begin{align*}
		E\left(m_c\right) &= E\left(m_a\right) \oplus \left(b^{e_b-e_a} \otimes E\left(m_b\right)\right) \text{and } e_c = e_a & \text{ if } e_a \leq e_b, \\
		E\left(m_c\right) &= E\left(m_b\right) \oplus \left(b^{e_a-e_b} \otimes E\left(m_a\right)\right) \text{and } e_c = e_b & \text{ if } e_a > e_b.
	\end{align*}
	% \begin{align*}
	% 	E\left(m_c\right) &= 
	% 	\begin{cases}
	% 		E\left(m_a\right) \oplus \left(b^{e_b-e_a} \otimes E\left(m_b\right)\right) & \text{if } e_a \leq e_b \\
	% 		E\left(m_b\right) \oplus \left(b^{e_a-e_b} \otimes E\left(m_a\right)\right) & \text{if } e_a > e_b
	% 	\end{cases} \\
	% 	% \text{ and }
	% 	e_c &= \min\left(e_a, e_b \right).	
	% 	% \begin{cases}
	% 	% 	e_a & \text{if } e_a \leq e_b \\
	% 	% 	e_b & \text{if } e_a > e_b
	% 	% \end{cases}.
	% \end{align*}
  \item[Scalar multiplication.]
    To compute $E\left(c\right) = E\left(ab\right)$, where $a$ and $E\left(b\right)$ are known (i.e., $m_a$ is not encrypted), we compute
    \begin{align*}
      E\left(m_c\right) &= m_a \otimes E\left(m_b\right),\\
      e_c &= e_a + e_b.
    \end{align*}
\end{description}

>>>>

\subsubsection{Key Generation}
We choose two large primes $p$ and $q$, and set $n = pq, \lambda = \mathrm{lcm}\left(p-1,q-1\right)$.
We then define $L\left(x\right)$ to be the largest integer $v$ greater than zero such that $x-1 \geq vn$.
Then we select an integer $g$ such that $\gcd\left(L\left(g^\lambda \bmod n^2\right), n\right) = 1$ and $0\leq g \leq n^2$.
We denote the public key as $(g,n)$ and the private key $(p,q)$.
\subsubsection{Encryption and Decryption}
The encryption function to encrypt a plaintext $m \in \mathbb{Z}_n$ given a public key $(g,n)$ is defined as
\begin{equation*}
	E\left(m\right) = g^m \cdot r^n \mod{n^2},
\end{equation*}
where $r$ is a random non-negative integer less than $n^2$.
The decryption function to decrypt a ciphertext $c \in \mathbb{Z}^*_{n^2}$ given a private key $(p,q)$ is defined as:
\begin{equation*}
	D\left(c\right) = L\left(c^\lambda \bmod n^2\right) \cdot \left(L\left(g^\lambda \bmod n^2\right)\right)^{-1} \mod n.
\end{equation*}
\subsubsection{Homomorphic Properties of the Paillier Cryptosystem}
The Paillier cryptosystem supports additive homomorphism as well as the multiplication of a plaintext scalar to an encrypted message. These homomorphic properties are used to add a plaintext scalar to a ciphertext, add two ciphertexts, and multiply a ciphertext by a plaintext scalar, respectively. They are defined as follows.

@@

We represent a floating-point (FP) number as a pair of two integers $(m,e)$ representing the mantissa and exponent of the FP number with respect to a base $b$. The mantissa $m$ is encrypted, while the exponent $e$ is unencrypted.
Let $a,b,c$ be FP numbers represented by the pairs $(m_a,e_a),(m_b,e_b),(m_c,e_c)$ respectively. Let $\oplus,\otimes$ represent the homomorphic operations which correspond to the addition and multiplication of integers in the Paillier cryptosystem, respectively. We define the corresponding FP number operations as follows:
\begin{description}
  \item[Addition.]
    To compute $E\left(c\right)=E\left(a+b\right)$ we compute
    % \begin{align*}
	% 	E\left(m_c\right) &= E\left(m_a\right) \oplus \left(b^{e_b-e_a} \otimes E\left(m_b\right)\right) \text{and } e_c = e_a & \text{ if } e_a \leq e_b, \\
	% 	E\left(m_c\right) &= E\left(m_b\right) \oplus \left(b^{e_a-e_b} \otimes E\left(m_a\right)\right) \text{and } e_c = e_b & \text{ if } e_a > e_b.
	% \end{align*}
	\begin{align*}
		E\left(m_c\right) &= 
		\begin{cases}
			E\left(m_a\right) \oplus \left(b^{e_b-e_a} \otimes E\left(m_b\right)\right) & \text{if } e_a \leq e_b \\
			E\left(m_b\right) \oplus \left(b^{e_a-e_b} \otimes E\left(m_a\right)\right) & \text{if } e_a > e_b
		\end{cases}, \\
		% \text{ and }
		e_c &= %\min\left(e_a, e_b \right).	
		\begin{cases}
			e_a & \text{if } e_a \leq e_b \\
			e_b & \text{if } e_a > e_b
		\end{cases}.
	\end{align*}
  \item[Scalar multiplication.]
    To compute $E\left(c\right) = E\left(ab\right)$, where $a$ and $E\left(b\right)$ are known (i.e., $m_a$ is not encrypted), we compute
    \begin{align*}
      E\left(m_c\right) &= m_a \otimes E\left(m_b\right),\\
      e_c &= e_a + e_b.
    \end{align*}
\end{description}

### 2019-02-14 pcsc-paper/parts/03_exponentialandlogarithmic.tex
Applied most of the reviewer comments
Still missing time benchmarking

\section{Logarithmic Function Approximation}
We now present our main contribution in this paper: a general method to improve the general accuracy of the logarithm approximation given by Khattri in~\cite{khattri_new_2009} for use in privacy-preserving calculation, provided that accuracy is only required for a specific range.

@@

Applying the five-point Gauss--Legendre quadrature rule to approximate $\int_{n}^{\alpha n + \alpha}{\frac{1}{x}\diff x}$ with $\alpha = 1/16$ using SageMath 8.3, we arrive at the approximation:
\begin{align}\label{eq:optimal_log_approximation}
	\begin{split}
		&\log\left(1+x\right) \\
		&\approx \frac{137x^5 + 26685x^4 + 617370x^3 - 6498630x^2 - 121239315x - 257804775}
		{30(x^5 + 405x^4 + 27210x^3 + 488810x^2 + 2536005x + 3122577)}\\
		&+ \log{16}.
	\end{split}
\end{align}

>>>>

\section{Logarithmic Function Approximation}
\label{sec:logarithm_approximation}
We now present our main contribution in this paper: a general method to improve the general accuracy of the logarithm approximation given by Khattri in~\cite{khattri_new_2009} for use in privacy-preserving calculation, provided that accuracy is only required for a specific range.

@@

Applying the five-point Gauss--Legendre quadrature rule to approximate $\int_{n}^{\alpha n + \alpha}{\frac{1}{t}\diff t}$ with $\alpha = 1/16$ using SageMath 8.3, we arrive at the approximation:
\begin{align}\label{eq:optimal_log_approximation}
	\begin{split}
		&\log\left(1+x\right) \\
		&\approx \frac{137x^5 + 26685x^4 + 617370x^3 - 6498630x^2 - 121239315x - 257804775}
		{30(x^5 + 405x^4 + 27210x^3 + 488810x^2 + 2536005x + 3122577)}\\
		&+ \log{16}.
	\end{split}
\end{align}

### 2019-02-14 pcsc-paper/parts/04_inversetrig.tex
3rd draft
apply 2nd round of comments
Applied most of the reviewer comments
Still missing time benchmarking

\section{Inverse Trigonometric Functions}
We will now show how a Gauss--Legendre quadrature approach can also be used for efficient privacy-preserving approximation of the inverse tangent function. Two well-known series representations of the inverse tangent function exist: the Maclaurin series expansion,
\begin{align} \label{eq:arctan_maclaurin}
	\arctan x = \sum_{n=0}^{\infty}{\frac{(-1)^nx^{2n+1}}{2n+1}}, \text{ for }|x|\leq 1,
\end{align}
and the following series expansion discovered by Euler in 1755, which converges more rapidly for all $x$~\cite{chien-lih_89.67_2005}.
\begin{align} \label{eq:arctan_euler}
	\arctan x = \sum_{n=0}^{\infty}
	{
	\frac
		{2^{2n}(n!)^2}
		{(2n+1)!}
	\frac
		{x^{2n+1}}
		{(1+x^2)^{n+1}}
	}.
\end{align}
We will compare the quadrature approach with the method of computing partial sums from each of these series.

@@

Figure~\ref{fig:arctan_error} shows that despite requiring a similar number of floating-point operations, the quadrature-based approximation in Equation~\ref{eq:arctan_quadrature} gives accurate approximations over a larger interval, making it a more viable candidate for efficient computation as seen in Figure~\ref{fig:arctan_comparison}.
\begin{figure}[!ht]
	\centering
	\begin{tikzpicture}
		\begin{semilogyaxis}[
			xmin=-40, xmax=40,
			ymin=1e-11, ymax=2,
			xlabel={$x$}, ylabel={Absolute error},
			grid=major,
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		\addplot[thick, ACMDarkBlue] table[x=x, y=gl] {plots/abs_err_arctan.txt};
		\addplot[thick, ACMOrange] table[x=x, y=euler] {plots/abs_err_arctan.txt};
		\legend{Gauss--Legendre quadrature, Euler's series}
		\end{semilogyaxis}
	\end{tikzpicture}%
	\caption{Comparison of the absolute error of the approximations for $\arctan x$ using quadrature approximation (Equation~\ref{eq:arctan_quadrature}) and the first five terms of Euler's series (Equation~\ref{eq:arctan_euler_partial})}
	\label{fig:arctan_error}
\end{figure}

>>>>

\section{Inverse Tangent Function Approximation}
\label{sec:arctan_approximation}
We will now show how a Gauss--Legendre quadrature approach can also be used for efficient privacy-preserving approximation of the inverse tangent function. Two well-known series representations of the inverse tangent function exist: the Maclaurin series expansion,
\begin{align} \label{eq:arctan_maclaurin}
	\arctan x = \sum_{n=0}^{\infty}{\frac{(-1)^nx^{2n+1}}{2n+1}}, \text{ for }|x|\leq 1,
\end{align}
and the following series expansion discovered by Euler in 1755, which converges more rapidly for all $x$~\cite{chien-lih_89.67_2005}.
\begin{align} \label{eq:arctan_euler}
	\arctan x = \sum_{n=0}^{\infty}
	{
	\frac
		{2^{2n}(n!)^2}
		{(2n+1)!}
	\frac
		{x^{2n+1}}
		{(1+x^2)^{n+1}}
	}.
\end{align}
We will compare the quadrature approach with the method of computing partial sums from each of these series.

@@

Figure~\ref{fig:arctan_error} shows that despite requiring a similar number of floating-point operations, the quadrature-based approximation in Equation~\ref{eq:arctan_quadrature} gives accurate approximations over a larger interval, making it a more viable candidate for efficient computation as seen in Figure~\ref{fig:arctan_comparison}.
\begin{figure}[!ht]
	\centering
	\begin{tikzpicture}
		\begin{semilogyaxis}[
			xmin=-40, xmax=40,
			ymin=1e-11, ymax=2,
			xlabel={$x$}, ylabel={Absolute error},
			grid=major,
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		\addplot[thick, ACMDarkBlue] table[x=x, y=gl] {plots/abs_err_arctan.txt};
		\addplot[thick, ACMOrange] table[x=x, y=euler] {plots/abs_err_arctan.txt};
		\legend{Gauss--Legendre quadrature, Partial sum of Euler's series}
		\end{semilogyaxis}
	\end{tikzpicture}%
	\caption{Comparison of the absolute error of the approximations for $\arctan x$ using quadrature approximation (Equation~\ref{eq:arctan_quadrature}) and the first five terms of Euler's series (Equation~\ref{eq:arctan_euler_partial})}
	\label{fig:arctan_error}
\end{figure}

### 2019-02-14 pcsc-paper/parts/05_testsforaccuracy.tex
3rd draft
apply 2nd round of comments
Applied most of the reviewer comments
Still missing time benchmarking

\section{Experiments on Accuracy}
We implemented privacy-preserving floating-point arithmetic using the system described in Section~\ref{sec:fp_arithmetic}, through the \texttt{python-paillier} library (\url{https://github.com/n1analytics/python-paillier}) running on Python~3.6.5.

@@

\begin{table}[ht]
	\caption{Comparison of inverse tangent approximation errors under Paillier}
	\label{tab:arctan_approximation}
	\begin{tabular}{lcc}
		\toprule
		Input & Partial sum of infinite series & Quadrature\\
		\midrule
		$\arctan -1$ & $6.564089 \times 10^{-1}$ & $4.629557 \times 10^{-1}$\\
		$\arctan -0.1$ & $4.523405 \times 10^{-3}$ & $1.451929 \times 10^{-3}$\\
		$\arctan -0.01$ & $4.665260 \times 10^{-6}$ & $1.472027 \times 10^{-6}$\\
		$\arctan 0$ & $0$ & $0$\\
		$\arctan 0.01$ & $4.665228 \times 10^{-6}$ & $1.472020 \times 10^{-6}$\\
		$\arctan 0.1$ & $4.523405 \times 10^{-3} $ & $1.451929 \times 10^{-3}$\\
		$\arctan 1$ & $6.564089 \times 10^{-1} $ & $4.629557 \times 10^{-1}$\\
	\bottomrule
\end{tabular}
\end{table}

>>>>

@@

\begin{table}[ht]
	\caption{Comparison of inverse tangent approximation errors under Paillier}
	\label{tab:arctan_approximation}
	\begin{tabular}{lcc}
		\toprule
		Input & Partial sum of Euler's series & Quadrature\\
		\midrule
		$\arctan -1$ & $6.564089 \times 10^{-1}$ & $4.629557 \times 10^{-1}$\\
		$\arctan -0.1$ & $4.523405 \times 10^{-3}$ & $1.451929 \times 10^{-3}$\\
		$\arctan -0.01$ & $4.665260 \times 10^{-6}$ & $1.472027 \times 10^{-6}$\\
		$\arctan 0$ & $0$ & $0$\\
		$\arctan 0.01$ & $4.665228 \times 10^{-6}$ & $1.472020 \times 10^{-6}$\\
		$\arctan 0.1$ & $4.523405 \times 10^{-3} $ & $1.451929 \times 10^{-3}$\\
		$\arctan 1$ & $6.564089 \times 10^{-1} $ & $4.629557 \times 10^{-1}$\\
	\bottomrule
\end{tabular}
\end{table}
